Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-24 RNARedPrint.tex

To formalize our algorithms, which iteratively merge evaluations of partial solutions, we extend the idea of atomic feature contributions, which are evaluated at sets of the form $\{x_1\mapsto v_1, \dots, x_d\mapsto v_p\}$. Let us call the latter object a \Def{partial sequence}; such an object will help specify partial knowledge on the sequence at some point of the algorithm. Easily, we can extend the definition of contributions $f$ to sets $\{x_1\mapsto v_1, \dots, x_p\mapsto v_p\}$, where $\{x_1\dots x_p\}$ is any super-set of $\dep(f)$ by ignoring the superfluous assignments $x\mapsto v$, where $x\not\in\dep(f)$.

@@

We compute the partition function (as specified by $\F$) by dynamic programming based
on a tree decomposition of $G_{\F}$, the dependency graph associated with $\F$.
Note that analogous algorithms could be easily derived to count valid sequences, or list sequences having minimal free-energy.
%This yields fixed parameter tractable (FPT) algorithms, based on the treewidth.

@@

Since our algorithms iterate over specific sets of sequence positions, we moreover define
the \Def{set $\partseqs(\Y)$ of all partial sequences determining the positions of $\Y=\subseteq\{1,\dots,n\}$} in all combinations of nucleotides $\{\Ab,\Cb,\Gb,\Ub\}$,
i.e.~for $\Y=\{y_1,\dots,y_{q}\}$,
\begin{align*}
\partseqs(\Y) = \{ & \{ y_i\mapsto v_i \mid i=1,\dots,q \} \\& | (v_1,\dots,v_{q}) \in \{\Ab,\Cb,\Gb,\Ub\}^{q} \}.
\end{align*}
%
We assume the following properties of the given cluster tree (reflecting $\F$):
\begin{itemize}
\item
$T$ is
connected and contains a dedicated node $r$, with
$\chi(r)=\varnothing$ and $\phi(r)=\varnothing$. If such a root does not exist, it can be added to the tree decomposition and connected to one node in each connected component of $T$;
\item
all edges in the tree decompostion are oriented towards this root;
\item
all sets $\difference{u}{v}$ are singleton: for any given
cluster tree, an equivalent (in term of treewidth) cluster tree can
always be obtained by inserting at most $\Theta(|\X|)$ additional
clusters.
\end{itemize}

>>>>

To formalize our algorithms, which iteratively merge evaluations of partial solutions, we extend the idea of atomic feature contributions, which are evaluated at sets of the form $\{x_1\mapsto v_1, \dots, x_d\mapsto v_p\}$. Let us call the latter object a \Def{partial sequence}. Such an object will help \revised{to} specify partial knowledge on the sequence at some point of the algorithm. Easily, we can extend the definition of contributions $f$ to sets $\{x_1\mapsto v_1, \dots, x_p\mapsto v_p\}$, where $\{x_1\dots x_p\}$ is any super-set of $\dep(f)$ by ignoring the superfluous assignments $x\mapsto v$, where $x\not\in\dep(f)$.

@@

We compute the partition function (as specified by $\F$) by dynamic programming based
on a tree decomposition of $G_{\F}$, the dependency graph associated with $\F$.
Note, that analogous algorithms could be easily derived to count valid sequences, or list sequences having minimal free-energy.
%This yields fixed parameter tractable (FPT) algorithms, based on the treewidth.

@@

Since our algorithms iterate over specific sets of sequence positions, we moreover define
the \Def{set $\partseqs(\Y)$ of all partial sequences determining the positions of $\Y=\subseteq\{1,\dots,n\}$} in all combinations of nucleotides $\{\Ab,\Cb,\Gb,\Ub\}$,
i.e.~for $\Y=\{y_1,\dots,y_{q}\}$,
\begin{align*}
\partseqs(\Y) = \{ & \{ y_i\mapsto v_i \mid i=1,\dots,q \} \\& | (v_1,\dots,v_{q}) \in \{\Ab,\Cb,\Gb,\Ub\}^{q} \}.
\end{align*}
%
We assume the following properties of the given cluster tree (reflecting $\F$):
\begin{itemize}
\item
$T$ is
connected and contains a dedicated node $r$, with
$\chi(r)=\varnothing$ and $\phi(r)=\varnothing$. If such a root does not exist, it can be added to the tree decomposition and connected to one node in each connected component of $T$;
\item
all edges in the tree decompostion are oriented towards this root;
\item
all sets $\difference{u}{v}$ are singleton: for any given
cluster tree, an equivalent (in term of treewidth) cluster tree can
always be obtained by inserting at most $\Theta(|\X|)$ additional
clusters.
\end{itemize}
\todo{$\Theta$ not defined, what is it?}
